Stochastic modeling of EGS using continuum-fracture approach and multilevel Monte Carlo

### 2019-07-10 paper.tex
Updates from Overleaf

\begin{tabular}{|l|c|c|c|}
\hline
 & rock & fractures & liquid\\\hline
Density $\varrho$ [kg.m${}^{-3}$] & 2700 & 2700 & 1000 \\
Hydraulic conductivity $k$ [m.s${}^{-1}$] & $10^{-9}$ \cite{sperl-trckova} & $10^{-3}$ & ---\\
Porosity $\vartheta$ [--] & $5\times10^{-3}$ \cite{intera} & $5\times10^{-2}$ & --- \\
Compressibility $\beta$ [Pa${}^{-1}$] & $2\times10^{-11}$ \cite{zisman} & $2\times10^{-11}$ & $5\times10^{-10}$ \\
Storativity $S$ [m${}^{-1}$] & $2\times10^{-7}$ & $2\times10^{-7}$ & --- \\
Cross-section $\delta$ [m] & --- & $10^{-6}$ & --- \\
\hline
Young modulus $G$ [Pa] & $50\times10^9$ \cite{ljunggren} & 50 & --- \\
Poisson's ratio $\nu$ [--] & 0.25 \cite{ljunggren} & 0.25 & --- \\
Biot's coefficient $\alpha$ [--] & 1 & 1 & --- \\
\hline
Heat capacity $c$ [J.kg${}^{-1}$.K${}^{-1}$] & 790 & 790 & 4000 \\
Thermal conductivity $\lambda$ [W.m${}^{-1}$.K${}^{-1}$] & 2.5 & 2.5 & 0.5 \\
\hline
\end{tabular}

Storativity is calculated as follows \cite{brace-et-al}:
\[ S = \varrho_l g(\beta_r + \vartheta\beta_l), \]
where $g$ is the constant of gravity, which gives approximately the same values both for rock and fractures.

@@

Number of wells: 2 \\
Depth: 5 km \\
Distance: 200 m \\
Radius: 15 cm \\
Length of open part of wells: 60 m \\

@@

On the surface of the artificial cylinders we shall extrapolate the value of piezometric head using the following idea: Given value $H_w$ of piezometric head at the well surface (with radius $r_w$) we assume that the piezometric head in the vicinity of the well is the function of the distance $d_w$ to the well axis:
\eq{\label{eq:H_log} H(\xx) = H_w\frac{\log d_w(|\xx|)}{\log r_w}. }
Then, on the surface of the cylinder at radius $r_c$ we impose the Robin condition in the Darcy flow equation:
\eq{\label{eq:H_robin} k\nabla H\cdot\nn = \sigma(H-H_w), }
where $\nn$ stands for the unit outward normal vector.
From \eqref{eq:H_log} and \eqref{eq:H_robin} we obtain the transition coefficient
\[ \sigma = \frac{k}{r_c}\left|\frac{\log r_c}{\log r_w}\right| \]
whose value depends on the hydraulic conductivity.

@@

In the first model scenario the fluid is injected to both wells in order to open the preexisting fractures.
The result is the modified cross-section of fractures.
For simplicity we stimulate both wells simultaneously which produces very similar result in the change of fracture cross-section as if the wells were stimulated one by one.
The liquid is injected for 1 day under the pressure 100 MPa (which corresponds to piezometric head $10^4$ m).

@@

Boundary conditions:
\begin{itemize}
\item piezometric head $10^4$ m on surface of wells (including the intersections with fractures)
\item piezometric head 0 m on lateral sides of the box (MOZNA BY MELO BYT I NAHORE A DOLE?)
\item no fluid flow through bottom and top part of the box and fracture tips
\item zero displacement on bottom part of the box
\item zero traction elsewhere
\end{itemize}

We use the Biot poroelasticity equations in the 3d continuum and 2d network of fractures. By dimension reduction of the Biot equations we obtain fluxes and forces which couple the rock with the lower dimensional fractures:
\[ \partial_t(Sp + \alpha\div\uu) - k\Delta p = 0,\qquad -\mu\div(\nabla\uu+\nabla\uu^\top) - \lambda\nabla\div\uu + \alpha\varrho g\nabla H = 0. \]
Here $p=H-x_3$ is the pressure head [m], $\uu$ is the displacement [m], $\mu$ and $\lambda$ are the Lamé parameters [Pa].
...

@@

The second scenario corresponds to operation of the EGS within 30 years after hydraulic stimulation performed according to Sec. \ref{sc:hm_model}.
The fluid is continuously injected into one well and pumped out of the second one under the pressure difference 20 MPa and with inlet temperature $15\ ^\circ$C.

@@

Boundary conditions:
\begin{itemize}
\item piezometric head $\pm10^3$ m on surface of wells ($+$ on the injection well, $-$ on the production well)
\item piezometric head 0 m on lateral sides of the box (MOZNA BY MELO BYT I NAHORE A DOLE?)
\item no flux on bottom and top part of the box and on fracture tips
\item temperature $15\ ^\circ$C on surface of injection well (ZDE BY SE MELA UDELAT APROXIMACE DIRICHLETA ROBINEM PODOBNE JAKO V PROUDENI Z NEJAKEHO ANALYTICKEHO RESENI)
\item initial temperature on lateral sides of the box where fluid flows into the box
\item no heat flux through bottom and top part of the box and fracture tips
\end{itemize}

>>>>

\begin{tabular}{|l|c|c|c|}
\hline
 & rock & fractures & liquid\\\hline
Density $\varrho$ [kg.m${}^{-3}$] & 2700 & 2700 & 1000 \\
Hydraulic conductivity $k$ [m.s${}^{-1}$] & $10^{-9}$ \cite{sperl-trckova} & $10^{-3}$ & ---\\
Porosity $\vartheta$ [--] & $5\times10^{-3}$ \cite{intera} & $5\times10^{-2}$ & --- \\
Compressibility $\beta$ [Pa${}^{-1}$] & $2\times10^{-11}$ \cite{zisman} & $2\times10^{-11}$ & $5\times10^{-10}$ \\
Storativity $S$ [m${}^{-1}$] & $2\times10^{-7}$ & $2\times10^{-7}$ & --- \\
Cross-section $\delta$ [m] & --- & $10^{-3}$ & --- \\
\hline
Young modulus $G$ [Pa] & $50\times10^9$ \cite{ljunggren} & 50 & --- \\
Poisson's ratio $\nu$ [--] & 0.25 \cite{ljunggren} & 0.25 & --- \\
Biot's coefficient $\alpha$ [--] & 1 & 1 & --- \\
\hline
Heat capacity $c$ [J.kg${}^{-1}$.K${}^{-1}$] & 790 & 790 & 4000 \\
Thermal conductivity $\lambda$ [W.m${}^{-1}$.K${}^{-1}$] & 2.5 & 2.5 & 0.5 \\
\hline
\end{tabular}

Storativity is calculated as follows \cite{brace-et-al}:
\[ S = \varrho_l g(\beta_r + \vartheta\beta_l), \]
where $g$ is the standard gravity [m.s${}^{-2}$], which gives approximately the same values both for rock and fractures.

@@

Number of wells: 2 \\
Depth: 5 km \\
Distance: 200 m \\
Well radius: 15 cm \\
Length of open part of wells: 60 m \\

@@

On the surface of the artificial cylinders we shall extrapolate the value of piezometric head as follows: Given value $H_w$ of piezometric head at the well surface (with radius $r_w$) we assume that the piezometric head in the vicinity of the well is the function of the distance $d_w$ to the well axis:
\eq{\label{eq:H_log} H(\xx) = H_w\frac{\log d_w(|\xx|)}{\log r_w}. }
Then, on the surface of the cylinder at radius $r_c$ we impose the Robin condition in the Darcy flow equation:
\eq{\label{eq:H_robin} k\nabla H\cdot\nn = \sigma(H-H_w), }
where $\nn$ stands for the unit outward normal vector.
From \eqref{eq:H_log} and \eqref{eq:H_robin} we obtain the transition coefficient
\[ \sigma = \frac{k}{r_c}\left|\frac{\log r_c}{\log r_w}\right| \]
whose value depends on the hydraulic conductivity.

@@

In the first model scenario the fluid is injected to both wells in order to open the preexisting fractures.
The result is the modified cross-section of fractures.
For simplicity we stimulate both wells simultaneously which produces very similar result in the change of fracture cross-section as if the wells were stimulated one by one.
The liquid is injected for 1 day under the pressure 10 MPa (which corresponds to piezometric head $10^3$ m).

@@

Boundary conditions:
\begin{itemize}
\item piezometric head $10^3$ m on surface of wells (including the intersections with fractures)
\item piezometric head 0 m on lateral sides of the box (MOZNA BY MELO BYT I NAHORE A DOLE?)
\item no fluid flow through bottom and top part of the box and fracture tips
\item zero displacement on bottom part of the box, zero displacement in the direction orthogonal to box sides
\item zero traction elsewhere
\end{itemize}

We use the Biot poroelasticity equations in the 3d continuum and 2d network of fractures. By dimension reduction of the Biot equations we obtain fluxes and forces which couple the rock with the lower dimensional fractures:
\[ \partial_t(Sp + \alpha\div\uu) - k\Delta p = 0,\qquad -\mu\div(\nabla\uu+\nabla\uu^\top) - \lambda\nabla\div\uu + \alpha\varrho g\nabla p = 0. \]
Here $p=H-x_3$ is the pressure head [m], $\uu$ is the displacement [m], $\mu$ and $\lambda$ are the Lamé parameters [Pa].
...

@@

The second scenario corresponds to operation of the EGS within 30 years after hydraulic stimulation performed according to Sec. \ref{sc:hm_model}.
The fluid is continuously injected into one well and pumped out of the second one under the pressure difference 2 MPa and with inlet temperature $15\ ^\circ$C.

@@

Boundary conditions:
\begin{itemize}
\item piezometric head $\pm10^2$ m on surface of wells ($+$ on the injection well, $-$ on the production well)
\item piezometric head 0 m on lateral sides of the box (MOZNA BY MELO BYT I NAHORE A DOLE?)
\item no flux on bottom and top part of the box and on fracture tips
\item temperature $15\ ^\circ$C on surface of injection well (ZDE BY SE MELA UDELAT APROXIMACE DIRICHLETA ROBINEM PODOBNE JAKO V PROUDENI Z NEJAKEHO ANALYTICKEHO RESENI)
\item initial temperature on lateral sides of the box where fluid flows into the box
\item no heat flux through bottom and top part of the box and fracture tips
\end{itemize}
